Opening up the directory on my computer for this file:
```{r, echo= TRUE}
setwd("C:/Users/Manuel/Desktop/Coursera")
datafile <- read.csv("./repdata/activity.csv")

head(datafile)
summary(datafile)

```

What is the mean of steps taken per day?

```{r, echo = TRUE}

#Getting the mean
#Fist: Add the number of steps per day
TotalSteps <- aggregate(steps ~ date, datafile, FUN = sum)
summary(TotalSteps)

#Make the histogram:
hist(TotalSteps$steps, xlab = " steps", ylab = "frequency", col = "red")

#Getting the mean:

StepsMean<- mean(TotalSteps$steps, na.rm = "TRUE")
StepsMedian <- median(TotalSteps$steps, na.rm = "TRUE")

#Printing the values for mean and median:
StepsMean
StepsMedian


```
Next: Making a time series plot

```{r , echo = TRUE}

#Getting the average number of steps in a 5min interval:
AvgSteps <- aggregate(steps ~ interval , datafile , FUN = mean)
summary(AvgSteps)
head(AvgSteps)


#Making the plot:
plot(AvgSteps$interval, AvgSteps$steps, type = "l", col = "red", xlab = "Interval" , ylab = "Avg Steps", main = "Average Number Steps Per 5-min Interval")

#Finding which 5 min interval has the most number of steps taken:
MostSteps <- which.max(AvgSteps$steps)
MostSteps   #This returns the number of the list, which is interval 835 , which has ~209 steps

#A more refined approach to getting the interval with the most steps is below:
maxInt <- AvgSteps[which.max(AvgSteps$steps),]
maxInt
```

Finding and Replacing the Missing Data

```{r, echo = TRUE}

#Total number of missing values is: 
sum(is.na(datafile[,1]))


#Replacing the missing data with Average in 5 min interval
missing <- !complete.cases(datafile)
REMNA <- datafile[missing == TRUE,]
REMNA[,1] <- AvgSteps$steps

summary(REMNA) #Shows no NA values

#Making a newFile with replaced data
datafile1 <- rbind(datafile[complete.cases(datafile),], REMNA)

#Getting the New Average with NA values filled in
NewSum <- aggregate(steps ~ date, datafile1, FUN = sum, na.rm = TRUE)

summary(NewSum)

#Making a histogram of total number of steps in new file
hist(NewSum$steps, xlab = " steps", col = "red")



```

Capturing the Diferrence between Weekend and Weekday activity
```{r, echo = TRUE}

#labeling Weekdays and weekends appropropiately
DayLabel <- function(date) {
  day <- weekdays(date)
  if (day %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday'))
      return ("weekeday")
  else if (day %in% c('Saturday', 'Sunday'))
      return ("weekend")
  else
      stop ("Invalid Date Format.")
}

datafile1$date <- as.Date(datafile1$date)
datafile1$day <- sapply(datafile1$date, FUN = DayLabel)

```

Plotting the Data
```{r, echo = TRUE}
Label_Days_Mean <- aggregate(steps ~ interval + day, datafile1,  mean)

head(Label_Days_Mean)

#Plotting the Data
library(lattice)
xyplot(steps ~ interval | day, data = Label_Day_Means, type = "l", xlab = "Interval", 
    ylab = "Number of steps", layout = c(1, 2))